\section*{Acknowledgments}\label{acknowledgments}
\thispagestyle{empty} % Remove page number from acknowledgments

There is no one more deserving of thanks than my very dedicated supervisor, Dr. Roby Austin (Saint Mary's University - Dept. of Astronomy and Physics).
I have had the great advantage of her endless support throughout my undergraduate degree and my masters.
I am not sure if it was my curiosity or hers that was so contagious, but her indulgence has resulted in so much encouragement and a wonderful picture of how science should work.
Thank you Roby, for your considerable and consistent efforts in correcting, suggesting, motivating, and inspiring.

I would also like to thank my committee members, Dr. Adam J. Sarty (Saint Mary's University - Dept. of Astronomy and Physics), Dr. Paul Muir (Saint Mary's University - Dept. of Mathematics and Computing Science), Dr. Pawan Lingras (Saint Mary's University - Dept. of Mathematics and Computing Science), and my external examiner Dr. Svetlana Barkanova (Acadia University - Dept. of Physics).
Their support over the past 3 years has been critical to completing this work.

My partner in crime and life, Alicia Beazley, has provided so much support when I needed it most.
Thank you for your understanding, reassurance, and reminders that I actually need to eat and sleep.

I appreciate so much my family's support.
I am very happy I have made them proud.

Many thanks to the Mathematics and Computing Science department which has provided a very valuable collection of minds that were a constant source of suggestions and support.
Finally, I would like to acknowledge that I cannot thank everyone who has helped of the last 3 years because there are so many of you, and for that I am very grateful.

