\thispagestyle{empty} % Remove page number from title page
\pagebreak
\thispagestyle{empty} % Remove page number from abstract
\begin{abstract}
\large Data Mining and Exploration of the Nuclear Science References

\large by Andrew Valencik

\normalsize
The Nuclear Science References (NSR) is a carefully curated bibliographic dataset focused on nuclear science literature.
A domain-specific search engine and supporting tools have been developed to aid and encourage the exploration of the NSR.
User queries are analyzed to form a series of filters to retrieve relevant NSR entries from a database.
The resulting information is presented in multiple views including lists, bar charts, and network graphs.
The network graph representations offer unique insights on collaborations centered around a given parameter such as a nuclide or group of authors.
The capability of clustering algorithms to expose trends within the dataset is demonstrated by clustering authors based on publication traits.
A vector space model based on the metadata provided in the NSR is used to recommend semantically similar NSR entries.
The completed work serves as both an example and a framework for future analysis of the NSR.

\large December 18th, 2015

\end{abstract}
%Data retrieved from user queries is used to form a graph data structure, enabling advanced visualizations and the use of various algorithms from graph theory.

\section*{Acknowledgments}\label{acknowledgments}
\thispagestyle{empty} % Remove page number from acknowledgments

There is no one more deserving of thanks than my very dedicated supervisor, Dr. Roby Austin (Saint Mary's University - Dept. of Astronomy and Physics).
I have had the great advantage of her endless support throughout my undergraduate degree and my masters.
I am not sure if it was my curiosity or hers that was so contagious, but her indulgence has resulted in so much encouragement and a wonderful picture of how science should work.
Thank you Roby, for your considerable and consistent efforts in correcting, suggesting, motivating, and inspiring.

I would also like to thank my committee members, Dr. Adam J. Sarty (Saint Mary's University - Dept. of Astronomy and Physics), Dr. Paul Muir (Saint Mary's University - Dept. of Mathematics and Computing Science), Dr. Pawan Lingras (Saint Mary's University - Dept. of Mathematics and Computing Science), and my external examiner Dr. Svetlana Barkanova (Acadia University - Dept. of Physics).
Their support over the past 3 years has been critical to completing this work.

My partner in crime and life, Alicia Beazley, has provided so much support when I needed it most.
Thank you for your understanding, reassurance, and reminders that I actually need to eat and sleep.

I appreciate so much my family's support.
I am very happy I have made them proud.

Many thanks to the Mathematics and Computing Science department which has provided a very valuable collection of minds that were a constant source of suggestions and support.
Finally, I would like to acknowledge that I cannot thank everyone who has helped of the last 3 years because there are so many of you, and for that I am very grateful.

\clearpage
\pagenumbering{arabic} % Arabic page numbers (and reset to 1)
\setcounter{page}{1}

