\documentclass[12pt]{article}

\usepackage{hyperref}
\usepackage[round]{natbib}
\bibliographystyle{plainnat}

\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}

\begin{document}

\title{Digital Signal Processing in Nuclear Physics with Field Programmable Gate Arrays}

\author{Andrew Valencik}

\date{\today}

\maketitle

%%%

\section{Field Programmable Gate Arrays}   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This section is about FPGAs
\\[20pt]

{\large\textbf{\cite{currentState}}}

The article is an industry produced review of FPGAs in the nuclear power sector. It has a good and thorough introduction to FPGA technology, including a discussion of hardware aspects such as CMOS, chip archetecture, different implementations (fuse, SRAM, antifuse), electrical and mechanical propertires, and the reliability of FPGAs.
\\
A good deal of the discussion focuses on the design and development of applications for FPGAs. The development process includes aspects from software design along with concerns and practices in hardware design. Finally, due to the nature of a nuclear power plant, an indepth analysis is given on the concerns of safety and risks associated with a FPGA system. This includes suggestions such as triple modular redudancy (TMR), which is covered in detail, using error detection and correction codes, and diverse methods of implementation.
\\[20pt]

{\large\textbf{\cite{Wirthlin}}}

The embedded applications, field reprogrammability, low non-reocurring engineering cost, and the short design cycle of FPGAs means they end up in a large range of environments. This study focuses on FPGAs in high radiation environments and draws largely off the knowledge base gained from the use of FPGA in space. 
\\
The defined bitstrem, flip-flops, registers of a FPGS are sensitive to heavy ion and proton single event upsets that might occur near a radiation detector. An expensive option to protect against single evenet upsets is to use triple modular redundancy. This involves duplicating logic components and comparing the results after various operations to protecting againt the failure of a logic unit. Another method is configuration scrubbing by reflashing the bitstream on the fly, to avoid the build up of failures. It is likely that the efforts to migate radiation effects in space will be useful in particle physics experiments as well. There is an ungoing effort to used FPGAs in the liquird argon calorimeter at ATLAS. 
\\[20pt]

{\large\textbf{\cite{Cromaz}}}

For the GRETINA spectrometer a multichannel digital signal processing board has been custom built using FPGAs. The board has 12-bit resolution and is capable of digitizing continuously at 100MHz. The FPGA on the board derives an energy, leading-edge time, and constant-fraction time. The board discussed was a prototype to the actual boards which use 10 channels and 14-bit ADCs which should improve issues with differential non-linearity. The FPGA used in the production board is an commercially available one, the Xilinx Spartan XC3S5000.
\\[20pt]


{\large\textbf{\cite{Ugur}}}

One of the most important aspects of particle identification experiments is the digitisation of time, amplitude and charge data from detectors. These conversions are mostly undertaken with Application Specific Integrated Circuits (ASICs). However, recent developments in Field Pro- grammable Gate Array (FPGA) technology allow us to use commercial electronic components for the required Front-End Electronics (FEE) and to do the digitisation in the FPGA. It is possible to do Time-of-Flight (ToF), Time-over-Threshold (ToT), amplitude and charge measurements with converters implemented in FPGA. 
\\[20pt]


\section{Digital Signal Processing}   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


{\large\textbf{\cite{Menaa2011512}}}

A significant advantage of digital signal processing over analog processing is the ability to apply a range of pulse shaping filters to the input signal. Menaa et al explore the use of cusp, 1/f, and pseudo-gaussian, pulse shaping in a FPGA implementation and compare it to the established CANBERRA trapezoidal shaper. The study concludes that for high throughput applications the trapezoidal filter has the best energy resolution. However in low noise germanium applications the cusp filter is ideal.
The article goes into little detail about the FPGA implementation. It nonetheless serves as a useful resource on different pulse shaping techniques and their best applications.
\\[20pt]


{\large\textbf{\cite{Suarez2008276}}}
Digital Signal Processing allows for complex pulse shape analysis, but if the algorithms are of sufficient complexity to require offline processing this produces new problems of handling and recording such data. A real time digital signal processing method is presented that accomplishes self-calibrating pulse shape analysis. Both fitting and parametric pulse analysis is discussed, but the focus is on the latter. Parametric analysis suffers some information loss with the added benefit of computational simplicityallowing it to be done live on FPGAs.
The article present an education view of the process of pulse shape discrimination. The work is conducted on a data collection board that features 14-bit ADCs and a FPGA for trapezoidal filtering and looking at pileup information. The algorithms of interest are walked through in detail.
\\[20pt]


{\large\textbf{\cite{Abbene2013124}}}
A commercial digitizer is used with a custom firmware for online pulse shape and height analysis. Both X-ray and gamma-ray spectra measurements are examined. A discussion of the application of paralyzable dead times to fast and slow pulse shape analysis is given. A comparison of digital signal processing and analog processing is given in regards to energy resolution as a function of input counting rate. This demonstrates the strong advantages the digital approach has over analog but also the effects of online shaping techniques at high throughput.
Thepulse height and shape analysis uses and describes the single delay line technique.Pile up rejection is performed by the use of a simple threashold. Dead times and the appropriate corrections are discussed in detail.
\\[20pt]


\section{Nuclear Physics}   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\large\textbf{\cite{Mayer}}}

The orientation dependence of the backscattered yield of 1.0-MeV helium ions has been used to investigate the lattice characteristics of silicon and germanium implanted at room temperature with 40-keV heavy ions (Ga, As, Sb, In, P). 
\\[20pt]


\section{Detectors}   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\large\textbf{\cite{Leviner201466}}}

The first N-type segmented Germanium detector with 85\% $^{76}\mbox{Ge}$ enrichment is characterized and evaluated for use in the Majorana collaboration. The segmented enriched germainium array (SEGA) is applicable to neutrinoless double beta-decay because of the strong energy resolution performance near 2039keV. The paper examines the cross-talk, electronic noise, and resolution of the detector. It concludes that SEGA is a viable detector for the search for neutrinoless double beta-decay.
\\
Neutrinoless double beta-decay has yet to be observed, but would show that neutrinos are mahorana particles, meaning they are their own antiparticle. These interactions have incredibly long halflives and are thus very rare.
\\[20pt]


{\large\textbf{\cite{Sangsingkeow2003183}}}

An older paper that summarizes the then recent advancements in high purity germanium detectors. Manufactoring and fabrication methods are the focus of the improvements. Many advancements had been made in growing large Germanium crystals that allowed for many segment detectors. Large efficieny HPGe detectors are presented with FWHM resolutions of 2.4keV at 1.33MeV. Multi-element arrays of Germanium detectors are introduced including the now prevalent clover array.
\\
Monolithic segmentation is an improvement covered in detail. This technique also for further information gathering from the same germanium crystal. Photolithography enables the use of very closely packed segmented electrical contacts. Thus instead on one electrical output signal from a Germanium crystal, there can be many outputs. Along with reduction in Doppler broadening, the segmentation means the addition of position information useful in tracking the path of a radiation event in the crystal.
\\[20pt]


{\large\textbf{\cite{Eberth2008283}}}

A lengthy and detailed topical review of gamma-ray tracking with Germanium detectors over the last 50 years. Gamma-ray spectroscopy is the focus, as it has been a primarily method for probing the structure of the atomic nucleus over the decades. The first improvements where in Germanium purity and volume. Segmentation by separating electrical contacts to the crystals came as a later improvement. Advances in signal processing allowed for more complex pulse shape analysis to be done on the increase in information gathered from segmentation. The European AGATA gamma-ray tracking system is the highlighted equipment of the review.
\\
Various components and their history are discussed and compared. Bismuth germanate and NaI(Ti) supression shields are discussed. A strong theoretical knowledge base is constructed for many properties of germanium semiconductor detectors. This includes discussions from dead layers to the different ways of tiling a sphere for a detector setup.
\\[20pt]


{\large\textbf{\cite{Descovich2005535}}}
In-beam position measurements are checked againt Monte Carlo simulations for highly segmented icoaxial germanium detectors. The determined resolution is 2mm in three directions for analysis on $^{90}\mbox{Zr}$ gamma transitions. The study examines the departure of experimental results from simulated data and break down the contributions of position resolution into the various factors of the detector.
The importance of position measurments is in enabling the reconstruction of the scattering path withing the detecting volume. In order to calculate the position of the first gamma interaction a Doppler correction must be made. The identified area to improve was in selecting a zero time for pulse shape analysis. This work is therefore very sensitive to improvements in timing information. Time and energy information were calculated using FPGAs on board.
\\[20pt]


{\large\textbf{\cite{Paschalis201344}}}
The Gamma-Ray Energy Tracking In-beam Nuclear Array, GRETINA, detector modules maximize solid angle coverage and tracking performance while minimizing costs in detector manufactoring. The energy and point position of each gamma interaction can be reconstructed with high precision. A detailed discussion on detector geometry is given to justify the design concepts of GRETINA. The detector modules are n-type high purity germanium crystals segmented 36 times. The testing and acceptance process is important is achieving the energy resolution of 2.2keV at 1333keV. The digital signal processing has 10 channel inputs, each leading to a 14-bit ADC with a 100MHz sampling rate. This information is then fed to a Xilinix FPGA which uses a trapezoidal filter to calculate energy, and leading edge and constant fraction discrimination algorithms for timing information.
\\[20pt]


{\large\textbf{\cite{Palit201290}}}
A compton supressed germanium clover detector array with a Pixie-16 commercial digital gamma finder digitizer has been implemented at TIFR-BARC. Gamma-gamma coincidence measurements are made using logic operations on the four clover output. The discussion include point of improvement from moving to the digital system from the previous analog electronics. Noteably, the previous system had limitations due to heat build up causing a gain drift. The experimets performed required a constant gain over a period of several days. The presented digital data aquisition system has improved data recording capabilities at INGA by a factor of five.
\\[20pt]


%%%

\bibliography{citationsAPSC.bib}

\end{document}
